Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-03-30 Bubble-deformation.tex
incorporating new results

%% Ohta
{\color{blue} Next, we examine the validation of our computational method
intended for the bubble deformation in simple linear shear flow reported by
Muller-Fischer et al.~\cite{MulTobDreFisWin08}.  Muller-Fischer et
al.~\cite{MulTobDreFisWin08} experimentally inquired into the bubble
deformation under the condition of $Re \approx 0$. In this study, using the
computational domain with $\lwh{12}{6}{6}$ and the finer grid size with $\Delta
x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$, computations were
performed for the condition of $Ca = 0.96$ and  1.63 with $Re \approx 0$ ($Re =
5.0 \times 10^{-4}$),  $\lambda = 1.2 \times 10^{-3}$ and  $\eta \leq 1.2 \times 10^{-6}$
according to experimental conditions by Muller-Fischer et al.
\cite{MulTobDreFisWin08}.  Comparison of our numerical results and previous
experimental results (\cite{MulTobDreFisWin08}) are tabulated in
Table~\ref{tab:DeComparisonRe=0}.  Additionally, in
Table~\ref{tab:DeComparisonRe=0}, we also list experimental results with the
condition of $Re \approx 0$ and $\lambda \approx \eta \approx 0$ by Rust and
Manga (\cite{RusMan02}). These experimental values by Rust and Manga
(\cite{RusMan02}) were obtained from the graph showing the relation of $D$ vs
$Re$ (\cite{RusMan02}).  As is clear from Table~\ref{tab:DeComparisonRe=0}, our
numerical results predicted larger values of $D$ than experimental ones
reported by Muller-Fischer et al.~\cite{MulTobDreFisWin08}. Nevertheless, we
found that our numerical results are very close to the experimental results by
Rust and Manga (\cite{RusMan02}), which emphasizes the intrinsic difficulties
associated with experimental investigations of bubble dynamics, even in simple
linear shear flow.  These comparisons suggest that our computational method is
effective and robust at reproducing bubble dynamics in simple linear shear
flow.}

%%
\begin{table}[tbh]
{\color{blue}
\caption{Comparison of  {\color{red} $D$} for a bubble as a function 
         of $Ca$ ($Re \approx 0$, $\lambda \approx \eta \approx 0$). In all cases, the domain 
         size is $\lwh{12}{6}{6}$.
         CLSVOF computations used a two-level AMR computational domain 
         with a level-0 discretization $\Delta x^{\ell=0} = \Delta y^{\ell=0} 
         = \Delta z^{\ell=0} = R/8$ and a level-1 discretization
         $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.}
 }
\label{tab:DeComparisonRe=0}
%\footnotesize
\center
\begin{tabular}{ c  c  c  c  c }
\hline
\hline
Capillary number                                                                 & 0.96   &  1.63       \\
{$D$}(Muller-Fischer et al. (\cite{MulTobDreFisWin08}))   & 0.37   &   0.58     \\
{$D$}(Rust and Manga (\cite{RusMan02}))    & $0.71\pm 0.05$   &   $0.81\pm 0.02$     \\
{$D$}(Our study)                                              & 0.63   &   0.76     \\
\hline
\hline
\end{tabular}
\end{table}
%%

@@

The results in Table~\ref{tab:DomComparison} suggest that drop deformation is
promoted when we use a domain size with $W=4R$.  In contrast, the drop does not
break up and becomes stable with a deformed shape if we set $L$ large enough
and $W \geq 6R$ and $H \geq 6R$.  Since the value of {\color{red} $D$} for the domain size
$\lwh{8}{6}{6}$ differs by only $1.3\%$ with the value for the domain size of
$\lwh{8}{16}{16}$, in the results that follow we set the width to $W=6R$ and the
height to $H=6R$ to minimize the number of computational grid nodes along those
directions.  To determine the critical Reynolds number $Re_{\rm c}$ (with $Ca=0.3$),
we consider a domain size of $\lwh{24}{6}{6}$, and we find that the drop
reaches a stable state with deformation parameter {\color{red} $D$}=0.549 for $Re=1.0$,
while a value of $Re=1.1$ leads to drop breakup.
% 
%\begin{comment} 
\begin{figure}%[h!]
  \centering
  \includegraphics[width=\textwidth]{DeEvolution}
  \caption{Time evolution of the deformation parameter {\color{red} $D$} versus 
   a dimensionless
   time $T = \mathit{\Gamma} t$ for a bubble, obtained 
   with two grid systems at different resolutions. Both systems use a 
   two-level AMR grid with a uniform 
   finest-level grid size $\Delta x^{\ell=1} = \Delta y^{\ell=1} = 
   \Delta z ^{\ell=1}$. The evolution of {\color{red} $D$} using the first system, where
   $\Delta x^{\ell=1} = R/16$, is shown in black and it predicts that 
   bubble breakup occurs at a Reynolds number $Re=35$. 
   Results with the second system, shown in red, use $\Delta x^{\ell=1} 
   = R/24$, and they predict that breakup occurs at $Re=34$.  The capillary
   number corresponding to these results is $Ca=1.0$.
   ($\lambda = 1.2 \times 10^{-3}$, $\eta < 1.0 \times 10^{-3}$)}
  \label{fig:DeEvolution}
\end{figure}
%\end{comment} 
%

The numerical results presented in this and the previous section used a
finest-level grid size set to $\Delta x^{\ell=1}(= \Delta y^{\ell=1}= \Delta
z^{\ell=1})=R/16$. To verify the adequacy of this grid resolution, we present
grid refinement results for a bubble breakup simulation with $Ca=1.0$, which
corresponds to the most deformable and stretchable bubble case considered in
our numerical studies. We use two different grid systems, one with $\Delta
x^{\ell=1}=R/16$ and another with $\Delta x^{\ell=1}=R/24$, to determine
$Re_{\rm c}$.  Figure~\ref{fig:DeEvolution} shows the time evolution of the
deformation parameter {\color{red} $D$} over time for the two grid systems; the $x$-axis is
a dimensionless time defined by $T=\mathit{\Gamma} t$ and the $y$-axis is {\color{red} $D$}.
The results show that bubble breakup occurs at $Re_{\rm c}$ = 35 for the grid system
with $\Delta x^{\ell=1}=R/16$, while the finer computational grid with $\Delta
x^{\ell=1}=R/24$ predicts a critical value of $Re_{\rm c}$ = 34.  Note that although
the time evolution of {\color{red} $D$} for the two grid systems is consistent for both
resolutions (the predicted critical Reynolds numbers differ by $\sim 3\%$), the
computational time using the grid system with $\Delta x^{\ell=1}=R/24$ was more
than 6 times longer than the one based on the coarser system with $\Delta
x^{\ell=1}=R/16$.  We remark that in our search for $Re_{\rm c}$, we considered a
wide range of values of $Ca$ and we found it necessary to use a large $L$
($\sim 24R$) since for certain shear flows the bubble can stretch significantly
without breaking up.  Nevertheless, for the conditions presented in this
section, the results indicate that our numerical approach, even with a
finest-level resolution set to $\Delta x^{\ell=1}=R/16$, is capable of
accurately reproducing bubble deformation and breakup without sacrificing any
essential dynamical features.
{\color{red} Additional grid refinements may be desirable to perform further 
accuracy evaluations for bubble deformation and breakup.}

>>>>

%% Ohta
{\color{blue} Next, we examine the validation of our computational method in which
we compare with the ``bubble deformation in simple linear shear flow'' 
results reported by
Muller-Fischer et al.~\cite{MulTobDreFisWin08}.  Muller-Fischer et
al.~\cite{MulTobDreFisWin08} experimentally inquired into the bubble
deformation under the condition of $Re \approx 0$. In our study, using the
computational domain with $\lwh{12}{6}{6}$ and the finer grid sizes with 
(i)
$\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$, and
(ii)
$\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/24$, 
computations were
performed for the conditions of $Ca = 0.96$ and $Ca = 1.63$ 
with $Re \approx 0$ ($Re =
5.0 \times 10^{-4}$),  
$\lambda = 1.2 \times 10^{-3}$ and  $\eta \leq 1.2 \times 10^{-6}$.
The parameters that we have prescribed are consistent with the 
experimental conditions by Muller-Fischer et al.
\cite{MulTobDreFisWin08}.  
Comparisons of our numerical results and previous
experimental results (\cite{MulTobDreFisWin08}) are tabulated in
Table~\ref{tab:DeComparisonRe=0}.  
Additionally, in
Table~\ref{tab:DeComparisonRe=0}, we also list experimental results with the
condition of $Re \approx 0$ and $\lambda \approx \eta \approx 0$ by Rust and
Manga (\cite{RusMan02}). These experimental values by Rust and Manga
(\cite{RusMan02}) were obtained from the graph showing the relation of $D$ vs
$Re$ (\cite{RusMan02}).  As is clear from Table~\ref{tab:DeComparisonRe=0}, our
numerical results predicted larger values of $D$ than experimental ones
reported by Muller-Fischer et al.~\cite{MulTobDreFisWin08}. Nevertheless, we
found that our numerical results are very close to the experimental results by
Rust and Manga (\cite{RusMan02}), which emphasizes the intrinsic difficulties
associated with experimental investigations of bubble dynamics, even in simple
linear shear flow.  These comparisons suggest that our computational method is
effective and robust at reproducing bubble dynamics in simple linear shear
flow.}

%%
\begin{table}[tbh]
{\color{blue}
\caption{Comparison of  {\color{red} $D$} for a bubble as a function 
         of $Ca$ ($Re \approx 0$, $\lambda \approx \eta \approx 0$). 
	 In all cases, the domain 
         size is $\lwh{12}{6}{6}$.
         CLSVOF computations used a two-level AMR computational domain 
         with a level-0 discretization $\Delta x^{\ell=0} = \Delta y^{\ell=0} 
         = \Delta z^{\ell=0} = R/8$ and a level-1 discretization
         $\Delta x^{\ell=1} = \Delta y^{\ell=1} = \Delta z^{\ell=1} = R/16$.}
 }
\label{tab:DeComparisonRe=0}
%\footnotesize
\center
\begin{tabular}{ c  c  c  c  c }
\hline
\hline
Capillary number & 0.96  & 1.63  \\
{$D$}(Muller-Fischer et al. (\cite{MulTobDreFisWin08}))   & 0.37   &   0.58     \\
{$D$}(Rust and Manga (\cite{RusMan02}))    & $0.71\pm 0.05$   &   $0.81\pm 0.02$     \\
{$D$}(Our study $\Delta x^{\ell=1}=R/16$)  & 0.63   &   0.76     \\
{$D$}(Our study $\Delta x^{\ell=1}=R/24$)  & 0.63   &   0.75     \\
\hline
\hline
\end{tabular}
\end{table}
%%

@@

The results in Table~\ref{tab:DomComparison} suggest that drop deformation is
promoted when we use a domain size with $W=4R$.  In contrast, the drop does not
break up and becomes stable with a deformed shape if we set $L$ large enough
and $W \geq 6R$ and $H \geq 6R$.  Since the value of {\color{red} $D$} for the domain size
$\lwh{8}{6}{6}$ differs by only $1.3\%$ with the value for the domain size of
$\lwh{8}{16}{16}$, in the results that follow we set the width to $W=6R$ and the
height to $H=6R$ to minimize the number of computational grid nodes along those
directions.  To determine the critical Reynolds number $Re_{\rm c}$ (with $Ca=0.3$),
we consider a domain size of $\lwh{24}{6}{6}$, and we find that the drop
reaches a stable state with deformation parameter {\color{red} $D$}=0.549 for $Re=1.0$,
while a value of $Re=1.1$ leads to drop breakup.
% 
%\begin{comment} 
\begin{figure}%[h!]
  \centering
%  \includegraphics[width=\textwidth]{DeEvolution}
  \includegraphics[width=\textwidth]{Fig3-Comparison-of-De-New}
  \caption{Time evolution of the deformation 
   parameter {\color{red} $D$} versus 
   dimensionless
   time $T = \mathit{\Gamma} t$ for a bubble, obtained 
   with three grid systems at different resolutions. All three resolutions 
   use a two-level AMR grid with an effective fine resolution  
   grid size $\Delta x^{\ell=1} = \Delta y^{\ell=1} = 
   \Delta z ^{\ell=1}$. 
   The evolution of {\color{red} $D$} using the first system, where
   $\Delta x^{\ell=1} = R/8$, is shown in blue and it predicts that 
   bubble breakup occurs at a Reynolds number $Re=33$.
   The evolution of {\color{red} $D$} using the second system, where
   $\Delta x^{\ell=1} = R/16$, is shown in black and it predicts that 
   bubble breakup occurs at a Reynolds number $Re=35$.
   Results with the third system, shown in red, use $\Delta x^{\ell=1} 
   = R/24$, and they predict that breakup occurs at $Re=34$.  The capillary
   number corresponding to these results is $Ca=1.0$.
   ($\lambda = 1.2 \times 10^{-3}$, $\eta < 1.0 \times 10^{-3}$)}
  \label{fig:DeEvolution}
\end{figure}
%\end{comment}
%

The numerical results presented in this and the previous section used a
finest-level grid size set to $\Delta x^{\ell=1}(= \Delta y^{\ell=1}= \Delta
z^{\ell=1})=R/16$. To verify the adequacy of this grid resolution, we present
grid refinement results for a bubble breakup simulation with $Ca=1.0$, which
corresponds to the most deformable and stretchable bubble case considered in
our numerical studies. We use three different grid systems, 
(i) $\Delta x^{\ell=1}=R/8$,
(ii) $\Delta x^{\ell=1}=R/16$,
and 
(iii) $\Delta x^{\ell=1}=R/24$, in order to determine
$Re_{\rm c}$.  Figure~\ref{fig:DeEvolution} shows the time evolution of the
deformation parameter {\color{red} $D$} over time for the three grid systems; 
the $x$-axis is
a dimensionless time defined by $T=\mathit{\Gamma} t$ 
and the $y$-axis is {\color{red} $D$}.
The results show that bubble breakup occurs at $Re_{\rm c}$ = 35 for the grid system
with $\Delta x^{\ell=1}=R/16$, while the finer computational grid with $\Delta
x^{\ell=1}=R/24$ predicts a critical value of $Re_{\rm c}$ = 34.  
For the $\Delta x^{\ell=1}=R/8$ results, it is clear that the $R/8$ resolution is too
coarse in order to capture the proper break-up time, albeit the critical
Reynolds' number, $Re_{\rm c}$ = 33, was still close to the finer grid resolution
cases.
Note that although
the time evolution of {\color{red} $D$} for the two finer resolution
systems ($R/16$ and $R/24$) is consistent between the two,
(the predicted critical Reynolds numbers differ by $\sim 3\%$), the
computational time using the grid system with $\Delta x^{\ell=1}=R/24$ was more
than 6 times longer than the one based on the coarser system with $\Delta
x^{\ell=1}=R/16$.  We remark that in our search for $Re_{\rm c}$, we considered a
wide range of values of $Ca$ and we found it necessary to use a large $L$
($\sim 24R$) since for certain shear flows the bubble can stretch significantly
without breaking up.  Nevertheless, for the conditions presented in this
section, the results indicate that our numerical approach, even with a
finest-level resolution set to $\Delta x^{\ell=1}=R/16$, is capable of
accurately reproducing bubble deformation and breakup without sacrificing any
essential dynamical features.
{\color{red} Additional grid refinements may be desirable to perform further 
accuracy evaluations for bubble deformation and breakup.}
